---
title: "Reproducible Research Project 01"
author: "David Junor"
date: "August 6, 2016"
output: 
        md_document:
                variant: markdown_github
---

```{r setup, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reproducible Research
##Project 01

```{r Load libraries}
library(ggplot2)
library(Hmisc)
```
###1. Load the data from source file and pre-process
```{r}
# Read activity file into activity
activity <- read.csv(file="C:/Users/Dave/Documents/DataScienceSpecialization/ReproducibleResearch/Project01/Working/Data/activity.csv")

# Pre-process data
activity$date <- as.Date(activity$date)
```
###2. What is the mean total number of steps taken each day?
```{r}
# Calculate steps per day and create histogram
stepsDay <- tapply(activity$steps,activity$date,sum,na.rm=TRUE)
qplot(stepsDay,xlab = "Total steps/day",ylab = "Frequency (binwidth = 1000)",binwidth=1000)
meanTotalSteps <- mean(stepsDay,na.rm = TRUE)
medianTotalSteps <- median(stepsDay,na.rm = TRUE)
```
The mean number of steps per day is `r meanTotalSteps`.
The median number of steps per day is `r medianTotalSteps`.

###3.  What is the average daily activity pattern?
```{r}
# Create time series plot
average <- aggregate(x=list(meanSteps=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```
```{r}
ggplot(data = average,aes(x = interval,y = meanSteps))+geom_line()+xlab("5 minutes")+ylab("Average steps")
```
```{r}
# Calculate which interval contains maximum number of steps
maxInterval <- average[which.max(average$meanSteps),]
```
The interval with the maximum number of steps is: `r maxInterval`.

###4.  Input missing values in data and analyze differences in activity patterns between weekdays and weekends.
```{r}
numMissingValues <- length(which(is.na(activity$steps)))
```
The number of missing values is: `r numMissingValues`.

```{r}
# Impute missing values
activityImputed <- activity
activityImputed$steps <- impute(activity$steps, fun=mean)
stepsByDayImputed <- tapply(activityImputed$steps, activityImputed$date, sum)
meanISteps <- mean(stepsByDayImputed)
medianISteps <- median(stepsByDayImputed)
```
```{r}
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency (binwidth = 1000', binwidth=1000)
```

The mean steps/day is `r meanISteps` and the median is `r medianISteps`.

###5. Difference in activity patterns between weekdays and weekends?
```{r}
activityImputed$dateType <-  ifelse(as.POSIXlt(activityImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
# Make time series plots
averagedActivityImputed <- aggregate(steps ~ interval + dateType, data=activityImputed, mean)
```
```{r}
ggplot(averagedActivityImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
